### 2020-09-23 LaTeX/sections/appendix.tex
edits on appendix

\section{Extended discussion on cleaning up data structures}

@@

\subsection{Ethereum's gas model}

DApps written in high-level programming languages are compiled and translated into a compact representation (called ‘bytecode’) to be further executed on the Ethereum virtual machine (EVM). When a transaction is executed, each opcode in the execution path accrues a fixed, pre-specified amount of gas. The function caller will pledge to pay a certain amount of ETH (typically quoted in units of \textbf{Gwei}, where \textit{wei} is the smallest transactional unit of ETH) per gas, and miners are free to chose to execute it or ignore it. The function caller is charged for exactly the amount the transaction costs and they cap the maximum they are willing to be charged (\textit{gas limit})---if the cap is too low to complete the execution, the miner keeps the gwei and reverts the state of the DApp (as if the function never ran).

In Ethereum, transactions are bundled into blocks. A miner can include as many transactions (typically preferring transactions that bid the highest for gas) that can fit under a pre-specified \texttt{block gas limit}, which is algorithmically adjusted for every block. As of the time of writting, the limit is around 11M gas.

\subsubsection{Gas refunds.} In order to reconstruct the current state of Ethereum's EVM, a node must obtain a copy of every variable change since the genesis block (or a more recent `checkpoint' that is universally agreed to). For this reason, stored variables persist for a long time and, at first glance, it seems pointless to free up variable storage (and unclear what `free up' even means). Once the current state of the EVM is established by a node, it can forget about every historical variable changes and only concern itself with the variables that have non-zero value (as a byte-string for non-integers) in the current state (uninitialized variables in Ethereum have the value zero by default). Therefore, freeing up variables will reduce the amount of state Ethereum nodes need to maintain going forward.

@@

Now let us put this theory into practice. In \cm, traders preload their account with ERC-20 tokens and/or ETH. \cm tracks what they are owed using a mapping called \texttt{totalBalance} and allow traders to withdraw their tokens at any time. However if a trader submits an order (\ie ask for their tokens), the tokens are committed and no available for withdrawal until the market closes (after which, the tokens are either transferred to the buyer or, if the trade is not executed, returned to the trader). Committed tokens are also tracked in a mapping called \texttt{unavailableBalance}. A buyer can request a token withdrawal up to their total balance subtracted by their unavailable balance.

As the DApp runs \texttt{closeMarket()}, it starts matching the best bids and to the best asks. As orders execute, \texttt{totalBalance} and \texttt{unavailableBalance} are updated. At a certain point, the bids and asks will stop matching in price. At this point, every order left in the order book cannot execute (because the priority queue sorts orders by price, and so orders deeper in the queue have worst prices than the order at the head of the queue). Therefore all remaining entries in \texttt{unavailableBalance} can be cleared.

>>>>

\section{Extended Discussion on Cleaning up Data Structures}

@@

\subsection{Ethereum's Gas Model}

DApps written in high-level programming languages are compiled and translated into a compact representation (called ‘bytecode’) to be further executed on the Ethereum virtual machine (EVM). When a transaction is executed, each opcode in the execution path accrues a fixed, pre-specified amount of gas. The function caller will pledge to pay a certain amount of ETH (typically quoted in units of \textbf{Gwei}, where \textit{wei} is the smallest transactional unit of ETH) per gas, and miners are free to choose to execute that transaction or ignore it. The function caller is charged for exactly the amount the transaction costs and they cap the maximum they are willing to be charged (\textit{gas limit})---if the cap is too low to complete the execution, the miner keeps the Gwei and reverts the state of the DApp (as if the function never ran).

In Ethereum, transactions are bundled into blocks. A miner can include as many transactions (typically preferring transactions that bid the highest for gas) that can fit under a pre-specified \texttt{block gas limit}, which is algorithmically adjusted for every block. As of the time of writing, the limit is around 11M gas.

\subsubsection{Gas Refunds.} In order to reconstruct the current state of Ethereum's EVM, a node must obtain a copy of every variable change since the genesis block (or a more recent `checkpoint' that is universally agreed to). For this reason, stored variables persist for a long time and, at first glance, it seems pointless to free up variable storage (and unclear what `free up' even means). Once the current state of the EVM is established by a node, it can forget about every historical variable changes and only concern itself with the variables that have non-zero value (as a byte-string for non-integers) in the current state (uninitialized variables in Ethereum have the value zero by default). Therefore, freeing up variables will reduce the amount of state Ethereum nodes need to maintain going forward.

@@

Now let us put this theory into practice. In \cm, traders preload their account with ERC-20 tokens and/or ETH. \cm tracks what they are owed using a mapping called \texttt{totalBalance} and allows traders to withdraw their tokens at any time. However if a trader submits an order (\ie ask for their tokens), the tokens are committed and not available for withdrawal until the market closes (after which, the tokens are either transferred to the buyer or, if the trade is not executed, returned to the trader). Committed tokens are also tracked in a mapping called \texttt{unavailableBalance}. Sellers can request a token withdrawal up to their total balance subtracted by their unavailable balance.

\textblue{As the DApp runs \texttt{closeMarket()}, it starts matching the best bids to the best asks. As orders execute, \texttt{totalBalance} and \texttt{unavailableBalance} are updated. At a certain point, the bids and asks will stop matching in price. At this point, every order left in the order book cannot execute (because the priority queue sorts orders by price, and so orders deeper in the queue have worst prices than the order at the head of the queue). Therefore all remaining entries in \texttt{unavailableBalance} can be cleared.}
